Details of specific functions

### 2024-11-08 inst/gitbook/03_specific-tutorials.Rmd
draft add

@@

### Disparity data is distance

If your disparity data is a distance matrix, you can use the option `is.dist = TRUE` in `dispRity` to make sure that all the operations done on your data take into account the fact that your disparity data has distance properties.
For example, if you bootstrap the data, this will automatically bootstrap both rows AND columns (i.e. so that the bootstrapped matrices are still distances).
This also improves speed on some calculations if you use [disparity metrics](#disparity-metrics) directly implemented in the package by avoiding recalculating distances (the full list can be seen in `?dispRity.metric` - they are usually the metrics with `dist` in their name).

### Disparity metric is distance

>>>>

### Bootstrapping dimensions

In some cases, you might also be interested in bootstrapping dimensions rather than observations.
I.e. bootstrapping the columns of a matrix rather than the rows.

It's pretty easy! By default, `boot.matrix` uses the option `boot.by = "rows"` which you can toggle to `boot.by = "columns"`

```{r, eval = TRUE}
## Bootstrapping the observations (default)
boot_obs <- boot.matrix(data = crown_stem, boot.by = "rows")

## Bootstrapping the columns rather than the rows
boot_dim <- boot.matrix(data = crown_stem, boot.by = "columns")
```

In these two examples, the first one `boot_obs` bootstraps the rows as showed before (default behaviour).
But the second one, `boot_dim` bootstraps the dimensions.
That means that for each

```{r, eval = TRUE}
## Measuring disparity and summarising
summary(dispRity(boot_obs, metric = mean))
summary(dispRity(boot_dim, metric = mean))
```

### Bootstrapping with fine tuned probabilities

```{r, eval = TRUE}
boot.matrix(data = BeckLee_mat50)
```

@@

### Disparity data is a distance

If your disparity data is a distance matrix, you can use the option `dist.data = TRUE` in `dispRity` to make sure that all the operations done on your data take into account the fact that your disparity data has distance properties.
For example, if you bootstrap the data, this will automatically bootstrap both rows AND columns (i.e. so that the bootstrapped matrices are still distances).
This also improves speed on some calculations if you use [disparity metrics](#disparity-metrics) directly implemented in the package by avoiding recalculating distances (the full list can be seen in `?dispRity.metric` - they are usually the metrics with `dist` in their name).

#### Subsets

dist.data = TRUE

#### Bootstrapping

boot.by = "dist"

### Disparity metric is a distance
